White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-11-14 papers/wdmerger_II/paper.tex
Update the resolution effect of the detonation time

We terminate the simulation at $t = 9$ seconds, as we have found that for this
particular setup, all of the significant burning has completed by this time. At
this point, the total energy on the domain has become positive, typically about
$10^{51}$ ergs, and has stopped increasing. This means that the system has become
unbound due to nuclear energy release, and that no further meaningful nuclear energy
generation is occurring. In our discussions below, it helps to have some other
anchoring points in time when significant events occur. We (arbitrarily) define the
collision time, $t_c$, as the first time when the maximum density on the grid reaches
$5 \times 10^6\ \text{g / cm}^3$. This is slightly larger than the initial maximum
density of the two WDs, and is reached only when the WDs meet at the center of the
domain. We define the detonation time, $t_d$, as the time when the temperature
first reaches $4 \times 10^9\ \text{K}$.

@@

Now let us examine the results for a collision of two 0.64 $\msolar$ WDs. We use
128 zones in the radial dimension, corresponding to a spatial resolution of 312.5 km.
Timesteps are limited only by the hydrodynamic stability constraint (and the plotfile
interval constraint). Resolutions in the range of a few hundred km have been a common
choice in earlier papers on white dwarf mergers and collisions. No adaptive mesh
refinement is used. For this configuration, the collision time is $t_c = 6.47\ \text{s}$,
and the detonation time is $t_d = 7.43\ \text{s}$. The peak density at the detonation time
is $2.3 \times 10^7\ \text{g / cm}^3$. The detonation occurs at the center of the domain
where the two WDs initially meet.

@@

The collision time is about $t_{\rm c} = 6.85\ \text{s}$
at high resolution. At low resolution, the simulation is slightly different because
of the distortion imposed by a rectangular grid representing initially spherical stars,
so the collision time is slightly later. Then, we restarted the run from this checkpoint
and reset the time to zero (such that $t_{\rm c} = 0$). The new simulation runs until about
$t = 2.15\ \text{s}$, so that the full 9 seconds are simulated. During this second phase
the detonation occurs. An estimate of the detonation time, $t_{\rm d}$, that serves well
is the time when the peak density on the grid first reaches $5 \times 10^9\ \text{K}$.
We measure the detonation time relative to $t_{\rm c}$. Approximately speaking, the later
$t_{\rm d}$ is, the more nickel is produced. As time passes, more white dwarf material
slams into the center, compressing and heating up the gas. Both effects are favorable
to production of nickel. So, a later $t_{\rm d}$ implies that more material can be converted
to nickel. Similarly, the speed of the detonation matters. A slower speed detonation prolongs
the collision event, allowing more gas to be processed into nickel.

>>>>

We terminate the simulation at $t = 9$ seconds, as we have found that for this
particular setup, all of the significant burning has completed by this time. At
this point, the total energy on the domain has become positive, typically about
$10^{51}$ ergs, and has stopped increasing. This means that the system has become
unbound due to nuclear energy release, and that no further meaningful nuclear energy
generation is occurring. In our discussions below, it helps to have some other
anchoring points in time when significant events occur. We (arbitrarily) define the
collision time, $t_{\rm c}$, as the first time when the maximum density on the grid reaches
$5 \times 10^6\ \text{g / cm}^3$. This is slightly larger than the initial maximum
density of the two WDs, and is reached only when the WDs meet at the center of the
domain. We define the detonation time, $t_{\rm d}$, as the time when the temperature
first reaches $4 \times 10^9\ \text{K}$.

@@

Now let us examine the results for a collision of two 0.64 $\msolar$ WDs. We use
128 zones in the radial dimension, corresponding to a spatial resolution of 312.5 km.
Timesteps are limited only by the hydrodynamic stability constraint (and the plotfile
interval constraint). Resolutions in the range of a few hundred km have been a common
choice in earlier papers on white dwarf mergers and collisions. No adaptive mesh
refinement is used. For this configuration, the collision time is $t_{\rm c} = 6.47\ \text{s}$,
and the detonation time is $t_{\rm d} = 7.43\ \text{s}$. The peak density at the detonation time
is $2.3 \times 10^7\ \text{g / cm}^3$. The detonation occurs at the center of the domain
where the two WDs initially meet.

@@

At high resolution the collision time is $t_{\rm c} = 6.41\ \text{s}$, which is similar to the
collision time at low resolution. However, the detonation time is around $t_{\rm d} = 6.62\ \text{s}$,
which is significantly earlier than at low resolution due to the high resolution simulations
being better able to resolve the point of contact between the WDs. The peak density crosses
$10^7\ \text{g / cm}^3$ significantly earlier, and this is what triggers the burning sufficient
to generate an earlier detonation. Approximately speaking, the later $t_{\rm d}$ is, the more
nickel is produced. As time passes, more white dwarf material slams into the center, compressing
and heating up the gas. Both effects are favorable to production of nickel. So, a later $t_{\rm d}$
implies that more material can be converted to nickel.
